---
title: "homework_08"
author: "Lamija Semic"
date: "`3/19/25`"
output: html_document
---

**1: Run sample code**

**a: Read in/generate data vector**
``` {r}
library(ggplot2) 
library(MASS) 

#z <- rnorm(n=3000, mean=0.2)
#z <- data.frame(1:3000, z)
#names(z) <- list("ID", "myVar")
#z <- z[z$myVar>0,]
#str(z)
#summary(z$myVar)

```

**b: Plot histogram**
```{r}
#p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  #geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
#print(p1)

```


**c: Add empirical density curve**
``` {r}
#p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
#print(p1)


```


**d: Get maximum likelihood parameters for normal**
```{r}
#normPars <- fitdistr(z$myVar,"normal")
#print(normPars)
#str(normPars)
#normPars$estimate["mean"] 

```


**e: Plot normal probability density**
```{r}
#meanML <- normPars$estimate["mean"]
#sdML <- normPars$estimate["sd"]

#xval <- seq(0,max(z$myVar),len=length(z$myVar))

# stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 #p1 + stat

```


**f: Plot exponential probability density**
```{r}
#expoPars <- fitdistr(z$myVar,"exponential")
#rateML <- expoPars$estimate["rate"]

#stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 #p1 + stat + stat2

```


**g: Plot uniform probability density**
```{r}
#stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 #p1 + stat + stat2 + stat3
```


**h: Plot gamma probability density**
```{r}
#gammaPars <- fitdistr(z$myVar,"gamma")
#shapeML <- gammaPars$estimate["shape"]
#rateML <- gammaPars$estimate["rate"]

#stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
# p1 + stat + stat2 + stat3 + stat4

```


**g: Plot beta probability density**
```{r}
#pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
 # geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
 # xlim(c(0,1)) +
 # geom_density(size=0.75,linetype="dotted")

#betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
#shape1ML <- betaPars$estimate["shape1"]
#shape2ML <- betaPars$estimate["shape2"]

#statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
#pSpecial + statSpecial
```

**2: Sample dataset**

**a:Read in data**
```{r}
z <- read.table("antcountydata.csv", header=TRUE, sep=",")
str(z)
names(z) <- list("ID", "myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
print(z)
summary(z)
```

**b: Plot histogram**
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

```

**c: Add empirical density curve**
``` {r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)


```

**d: Get maximum likelihood parameters for normal**
```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] 

```

**e: Plot normal probability density**
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat

```

**f: Plot exponential probability density**
```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2

```

**g: Plot uniform probability density**
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

**h: Plot gamma probability density**
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```

**g: Plot beta probability density**
```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```

**3: Best-fitting distribution**
```
The gamma density distribution is the most fitting for this sample data.

```

**4: New data set**
```{r}
library(ggplot2) 
library(MASS) 

z <- rnorm(n=1718, mean=0.2)
z <- data.frame(1:1718, z)
names(z) <- list("ID", "myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

```

**b: Plot histogram**
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

```

**d: Get maximum likelihood parameters for normal**
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

**h: Plot gamma probability density**
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

data <- data.frame(x = rnorm(1718))


p <- ggplot(data, aes(x = x)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.2, color = "grey60", fill = "cornsilk", size = 0.2) +
  stat_function(fun = dnorm, args = list(mean = mean(data$x), sd = sd(data$x)), colour = "red")

print(p)
```


**Original data histogram plot**
``` {r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

**Original data probability curve**
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```

The two histogram profiles differ, the original being skewed right, and the similated one representing a normal distribution. Therefore, it doesn't seem like the model is accurately simulating realistic data that matches that of the data set I imported. 
