---
title: "homework_06"
author: "Lamija Semic"
date: "`(2/19/25)`"
output: html_document
---
Study: insecticide exposure results in increased levels of octopamine, demonstrated by a normal distribution.


**1** - **3**
```{r}

treatment <- rnorm(n=30, mean=0, sd=1)
print(treatment)

control<- rnorm(n=30, mean=0, sd=1)
print(control)

octopamine_data <- c(treatment, control)
print(octopamine_data)

octopamine_data <- data.frame(octopamine_data)


```
Justifications:

- Sample sizes: 30, minimum sample size needed to make statistically statistically significant observations of data.
- Means: 0, mean for normal distribution.
- Variances: 1, sd for normal distribution. 



**4** - **5**
```{r}

one.way <- aov(treatment ~ control, data = octopamine_data)
summary(one.way)

```


**6**

```{r}

treatmenti <- rnorm(n=10, mean=0, sd=1)
print(treatment)

controli<- rnorm(n=10, mean=0, sd=1)
print(control)

octopamine_datai <- c(treatmenti, controli)
print(octopamine_datai)

octopamine_datai <- data.frame(octopamine_datai)

one.wayi <- aov(treatment ~ control, data = octopamine_datai)
summary(one.wayi)


for (i in octopamine_datai) {
  cat (runif(1), "\n")
}

for (i in octopamine_datai) {
  cat (runif(5), "\n")
}

for (i in octopamine_datai) {
  cat (runif(10), "\n")
}

for (i in octopamine_datai) {
  cat (runif(30), "\n")
}

```

5 samples is the smallest effective size needed to detect a significant pattern between the treatment and control.


**7**

```{r}


for (i in octopamine_data) {
  cat (runif(1), "\n")
}

for (i in octopamine_data) {
  cat (runif(5), "\n")
}

for (i in octopamine_data) {
  cat (runif(10), "\n")
}

for (i in octopamine_data) {
  cat (runif(30), "\n")
}

```
5 samples is the smallest effective size needed to detect a significant pattern between the treatment and control that were originally hypothesized.
